What if you drop that? Developmental changes in curiosity about physical interactions

### 2021-01-11 paper/paper2021.Rmd
CogSci skeleton to report online experiments

---
#title: "Modeling children's intrinsically-motivated curiosity about object interactions"
#title: "Drop it like it's hot: Modeling children's curiosity about physical interactions"
#title: "A drop in the bucket: Progress towards modeling children's curiosity about physical interactions"
title: "What if you drop that? Modeling children's curiosity about physical interactions"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA
    \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\ Department of Educational Psychology, 1025 W. Johnson Street \\ Madison, WI 53706 USA}

abstract: >
     Curiosity is a fundamental driver of human behavior, and yet because of its open-ended nature 
     and the wide variety of behaviors it inspires in different contexts, it is remarkably difficult 
     to study in a laboratory context. A promising approach to developing and testing theories of 
     curiosity is to instantiate them in artificial agents that are able to act and explore in a 
     simulated environment, and then compare the behavior of these agents to humans exploring the 
     same stimuli. Here we propose a new experimental paradigm for examining children's--and curious
     AI agents'--curiosity about objects' physical interactions, using a task that is both open-ended
     enough to allow room for curiosity, but also constrained enough to make detailed behavioral 
     comparisons. We compare the choices of 51 children (2-6 years of age) to those of agents implementing 
     different types of curiosity, and find broad exploration, but also systematic preferences ...

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
```

```{r, load-data}
tower = read.csv2("../analysis/human-data/curiobaby_drop-data - cool tower.csv", sep=',', header=T, stringsAsFactors=F)
raw = read.csv2("../analysis/human-data/curiobaby_drop-data - drop exp.csv", sep=',', header=T, stringsAsFactors=F)
hum = subset(raw, Exclude!="Y")

hum = hum[-which(hum$StimSet=="bc" & hum$DropChoice=="bowl and dumbbell"),] # chose two - exclude

s_tr <- hum %>% group_by(SID, Age_Group) %>% summarise(max=max(Trial)) %>% arrange(max)
age = table(s_tr$Age_Group)
```

@@

<!-- Curiosity is important... end with core question about what gives rise to curiosity. -->
Curiosity is a hallmark aspect of human intelligence. From infants exploring the objects in their environment to scientists exploring the frontiers of our solar system, humans are highly motivated to seek out new knowledge and experiences. However, although such exploratory behavior has long been recognized as a critical component of human learning [@james1983talks] and cognitive development [@Gopnik2009;@Piaget1952], formal theories that explain human curiosity and how it drives exploratory behavior have remained elusive [@Kidd2015]. Moreover, extant theories have rarely provided quantitatively precise enough predictions to be directly compared to empirical measurements of curiosity-driven behavior in humans.

@@

Our computational theory is inspired by classic descriptive theories that posited that curiosity-driven behavior was the consequence of attentional capture by novel stimuli [e.g., @Berlyne1954; @Fantz1964], as well as theories providing qualitative explanations for curiosity-driven behavior as the result of learners' preferences for stimuli that are "moderately discrepant" in relation to their current knowledge state, and thereby provide opportunities to learn [@Kinney1976]. Recent studies have provided empirical support for this basic idea: XX-old infants are sensitive to stimulus complexity and prefer to look at moderately complex stimuli rather than overly simple or overly complex stimuli; @Kidd2012; @Kidd2014].
According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957].
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.  
Our approach leverages theory and insights from reinforcement learning (RL), including the value of mechanisms that equip RL agents with "intrinsic motivation" — a drive to explore the state space even when rewards are sparse or absent, and favor actions with uncertain outcomes, because they may result in the discovery of new policies with high expected values [@Schmidhuber2010].
Our modeling approach is resonant with recent work that has instantiated such intrinsic-motivation mechanisms in robots to help them learn robust ways to predict physical events in the world world [@Oudeyer2007; @Oudeyer2013], although this work typically involves pretraining models on a separate physical prediction task before implementing curosity-driven learning, and has not yet been directly compared to human behavior in the same prediction tasks.

Our empirical study is motivated by a large body of prior work in developmental psychology investigating the development of knowledge about physical objects, their properties, and how they interact (CITE), thus making exploration of physical objects a natural choice of domain to explore the implications of our theory. However, the measures used in this literature have typically been coarse-grained with respect to infant's and children's physical knowledge, and thus do not specify or constrain which events might moderately discrepant from this knowledge. For instance, a typical measure of physical curiosity has been longer looking times to surprising events than to expected events (e.g., objects appearing to pass through a wall; @Stahl2015; @Baillargeon2007). Other work that has investigated exploratory behavior in children have used more granular measures, such as counting the number of functions discovered while playing with novel and complex toy [e.g., @Cook2011; @Bonawitz2012; @Gweon2014]. Moreover, these studies have yet to directly constrain theories that predict which functions children will discover, and in what sequence. What both of the standard looking-time and toy-exploration measures fail to capture when characterizing children's curiosity are the ways in which children actively learn by intervening on the world and observing the consequences of their actions [@Gopnik2009; @gureckis2012self]. To address these limitations, we developed a novel physical exploration task in which children choose which series of physical experiments to perform and observe the results of each in real time.

@@

(Full model description here, or just a teaser? Maybe also make the below a joint "Paradigm" section, then human behavior, model behavior and comparison)

# Experiment

@@

### Participants
Participants were `r length(unique(raw$SID))` children recruited from the Children's Discovery Museum of San Jose and Bing Nursery School.
Participant exclusions were made based on cases where i) the participant did not complete more than half of the study play session or ii) the parent did not consent for video recording of study. 
After exclusions, results from `r length(unique(hum$SID))` were analyzed, including `r age[1]` 2-year-olds, `r age[2]` 3-year-olds, `r age[3]` 4-year-olds, `r age[4]` 5-year-olds, and `r age[5]` 6-year-olds.\footnote{The collected sample differed from our planned sample size of 16 each of 3-, 4-, and 5-year-olds due to availability of participants (see preregistration: https://osf.io/37qvb/).}

@@

```{r object-sets}
shapes = c("bowl","cone","dumbbell","octahedron","pentagon","pipe","pyramid","torus","trig prism", "space")

a = c("pyramid", "torus", "trig prism")
b = c("cone", "octahedron", "pipe")
c = c("bowl", "dumbbell", "pentagon")

hum <- hum %>% 
  mutate(DropSet=ifelse(StimSet %in% c("ba","ca"), "A", ifelse(StimSet %in% c("ab","cb"), "B", "C")),
         TargetSet=ifelse(StimSet %in% c("ab","ac"), "A", ifelse(StimSet %in% c("ba","bc"), "B", "C")))
```

@@

The set of 9 objects were divided into 3 subsets (A = {pyramid, torus, triangular prism}, B = {cone, octahedron, pipe}, and C = {bowl, dumbbell, pentagon}). 
These subsets served as sets (A, B, C) of target or drop objects in six successive blocks of 2 trials, making a total of 12 trials per participant.
For example, in Order 1, in the first block of 2 trials, set A served as the target objects from which the child chose to drop on the target objects (set B).
The target:drop block sequence for Order 1 was A:B, B:C, C:A, A:C, B:A, C:B. The sequence for Order 2 was the reverse of Order 1: C:B, B:A, A:C, C:A, B:C, A:B. Participants were assigned to condition in counterbalanced order.

Target objects were placed in a circular bin (25 in diameter x 10 in height). 
The bin was divided with tape into sections of equal area, and one target object from the appropriate set was placed in the center of each third. 
Drop objects were presented to participants on a table at approximately eye level.

```{r fig1-sets, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.25, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Sets of 3D objects used for dropping and as targets."}
img <- png::readPNG("figs/drop_sets.png")
grid::grid.raster(img)
```

@@

After the parent provided informed consent, children were assigned to Order 1 or Order 2. Participants were assigned to condition in counterbalanced order.
We introduced children to a set of 3D-printed toy objects ("blocks"). 
The child played a game where they could pick which of three blocks to drop in a bin containing three other blocks, to see what happens. 
An example trial is illustrated in Figure 2.
We then swapped target/drop blocks, based on assigned order sequence, and asked the child to do the same selection and dropping a dozen times. 
Finally, we asked the child to build a "cool" tower with any of the toy blocks for about one minute.

Based on piloting, we estimated the activity would would only require five minutes to complete. 
In both conditions, a video camera was used to record the play session from an angle above the bin, to show child's block selection and drop location as well as child's completed tower. 
After child notified the researcher that they were finished building their tower, the session was completed and camera was turned off.

```{r fig2-task, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Example trial in which the participant chose to drop the pyramid from set A on the pentagonal prism in the target set (B)."}
img <- png::readPNG("figs/DropTaskFigure.png")
grid::grid.raster(img)
```

### Drop Coding Procedure

Each session's video was manually coded for drop object choice and target location per trial.
Drop choice was defined as the participant's selection of one of the three available drop set objects, to be dropped into the bin containing the set of the three potential target objects. 
The bin was divided into three wedge-shaped sections of equal area, demarcated with a thin black line.
One of the target set's objects was placed in the middle of each section of the bin. 
Target location was defined as either i) the object the dropped object collided with, ii) a collision with empty space inside the bin, or iii) falling outside the bin.
If the dropped object collided with empty space, we also recorded the target object it landed closest to. 
We also recorded events where the dropped object bounced from empty space or a target object then collided with other target object(s), including the number of collisions and identity of the target(s) hit.

Trials were excluded if the child i) dropped an object outside the bin, ii) touched target object(s) prior to a drop, iii) dropped an object during set change, iv) selected more than one object to drop, or iv) dropped (or threw) object from too high (>1.5 m). 
Trials were also excluded if the experimenter made a mistake (e.g., used the wrong stimulus set or repeated a trial more than twice).

@@

```{r chance-comparison}
hum$TargetSpace = ifelse(hum$DropLocation=="space", hum$WhichSpace, hum$DropLocation)

report_chisq <- function(X) {
  # e.g. X^2 (2, N = 88) = 2.1, p = .35
  pval = round(X$p.value, 3)
  if(pval==0) { 
    pval = ", p<.001" 
  } else {
    pval = paste0(", p=",pval)
  }
  return(paste0("($X^2$(", X$parameter, ", N=", sum(X$expected), ") = ", round(X$statistic,2), pval, ")"))
}

# see if children's choices (per subset) are random
drop_ch = table(hum$StimSet, hum$DropChoice)
# chisq.test(rbind(colSums(drop_ch[,a]), rep(sum(drop_ch[,a]/3), 3)) )

Xa_drop = chisq.test( colSums(drop_ch[,a]) ) # p<.001
Xb_drop = chisq.test( colSums(drop_ch[,b]) ) # p=.01
Xc_drop = chisq.test( colSums(drop_ch[,c]) ) # p=.056

target_ch = table(hum$StimSet, hum$DropLocation)

Xa_targ = chisq.test( colSums(target_ch[,a]) ) # n.s.
Xb_targ = chisq.test( colSums(target_ch[,b]) ) # n.s.
Xc_targ = chisq.test( colSums(target_ch[,c]) ) # n.s.

targetloc_ch = table(hum$StimSet, hum$TargetSpace)
Xa_tloc = chisq.test( colSums(targetloc_ch[,a]) ) # n.s.
Xb_tloc = chisq.test( colSums(targetloc_ch[,b]) ) # n.s.
Xc_tloc = chisq.test( colSums(targetloc_ch[,c]) ) # n.s.

# 'misses' by age
space_age = table(subset(hum, DropLocation=="space")$Age_Group)
all_age = table(hum$Age_Group)
#Xmiss_age = chisq.test(space_age, p=all_age / sum(all_age)) # n.s.
prop_space = sum(space_age) / sum(all_age)

age_s <- hum %>% 
  group_by(SID, Age_Exact) %>%
  summarise(miss=sum(DropLocation=="space"), N=n()) %>%
  mutate(prop_miss = miss / N)
age_miss = cor.test(age_s$prop_miss, as.numeric(age_s$Age_Exact))

# now check DropChoice x TargetLocation interactions
# maybe we do chisq.test on particular combinations? e.g., for DropSet A and TargetSet B?
dA = subset(hum, DropSet=="A" & TargetSet=="B")
dto = table(dA$DropChoice, dA$DropLocation)[,1:3]
#dto = table(hum$DropChoice, hum$DropLocation)[,c(1:4,6:8,10:11)] # drop 'outside' and 'space'
Xdto = chisq.test(dto)

```

We first examine children's choice of drop objects and target objects to determine if children were choosing randomly, or had consistent preferences for some objects. 
We used chi-square test of independence on participants' drop choices from each set of objects (A, B, C).
Participants' drop choices from set A significantly differed from chance `r report_chisq(Xa_drop)`, as did drop choices from set B `r report_chisq(Xb_drop)`.
Participants' drop choices from set C did not significantly differ from chance `r report_chisq(Xc_drop)`.
Table 1 shows a summary of participants' choice of drop objects per set.
It is apparent that from set A, participants preferentially chose to drop the torus rather than the triangular prism (trig prism) or pyramid. 
For set B, children more often avoided the octahedron, and instead chose the pipe or cone.
Set C showed more equal rates of drop choice, but the dumbbell was somewhat more popular than the pentagonal prism (pentagon).

What objects did children target with their drops?
First, we noted that children very often did not hit any target object: on `r round(prop_space, 2)*100`% of trials, participants' dropped over empty space.
It is not possible to determine whether children intentionally missed the available target objects, or whether their misses were errors due to noise in fine motor control.
If the misses were just due to poor motor control, one might expect the number of misses by age to decrease, as older children should have better motor control.
However, there was no significant correlation between children's individual proportion of "missed" targets and their age (`r cor_apa(age_miss, format="rmarkdown", print=F)`), which may suggest that dropping objects on empty space may be of interest to children of all ages, and not solely determined by a lack of fine motor control.
Thus, we first analyze drops where participants' dropped objects directly hit the target objects.
Chi-square tests of children's target collisions revealed that they did not significantly differ from chance for set A `r report_chisq(Xa_targ)`, B `r report_chisq(Xb_targ)`, or C `r report_chisq(Xc_targ)`.
Table 2 shows a summary of participants' choice of target objects per set.

Finally, we examine a more lenient analysis children's choice of target location (i.e., the identity of the closest target), regardless of whether there was a collision of drop and target objects.
Chi-square tests of children's target collisions once more found that they did not significantly differ from chance for set A `r report_chisq(Xa_tloc)`, B `r report_chisq(Xb_tloc)`, or C `r report_chisq(Xc_tloc)`.
Table 3 shows a summary of participants' choice of target locations per set.

In summary, children showed some consistent preferences in their choice of objects to drop from each set, but no systematic preference for targeting particular objects.
We now investigate whether there were interactions of drop objects and targets that children found particularly appealing. [ToDo: big chisq test?]

Finally, we analyze how exploratory children were in their responses: across the 12 trials, what proportion of the 9 available objects did each child utilize as drop objects? What proportion of the objects were targeted?

```{r table1-drop, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
get_table_string <- function(tableX, set) {
  tt = sort(colSums(tableX[,set]))
  tstr = paste(paste0(names(tt), " (", tt, "),"), collapse=' ')
  tstr = gsub('.{1}$', '', tstr)
  return(tstr)
}
dA = get_table_string(drop_ch, a)
dB = get_table_string(drop_ch, b)
dC = get_table_string(drop_ch, c)

tabl <- tribble(~Set, ~`Drop Object (N)`, 
                "A", dA,
                "B", dB,
                "C", dC)
t1 <- xtable::xtable(tabl, caption="Children's drop object choices by set.")
#tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
#                       caption = "This table prints across one column.")
print(t1, type="latex", comment = F, table.placement = "H")
```

```{r table2-target, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tA = get_table_string(target_ch, a)
tB = get_table_string(target_ch, b)
tC = get_table_string(target_ch, c)

tabl2 <- tribble(~Set, ~`Target Object (N)`, 
                "A", tA,
                "B", tB,
                "C", tC)
t2 <- xtable::xtable(tabl2, caption="Children's target object choices by set.")
print(t2, type="latex", comment = F, table.placement = "H")
```

```{r table3-targetloc, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tA = get_table_string(targetloc_ch, a)
tB = get_table_string(targetloc_ch, b)
tC = get_table_string(targetloc_ch, c)

tabl3 <- tribble(~Set, ~`Target Location (N)`, 
                "A", tA,
                "B", tB,
                "C", tC)
t3 <- xtable::xtable(tabl3, caption="Children's target location choices by set.")
print(t3, type="latex", comment = F, table.placement = "H")
```

@@

fig_drop <- drop_choice_plot(hum, shapes)
#print(fig_drop)
```

```{r target-hit, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.4, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Relative frequency of targeted object as a function of stimulus set."}
hum$Target = hum$DropLocation

target_hit_plot <- function(dat, ymax=.66, shapes_vec) {
  ch_cond <- dat %>% 
    filter(Target!="outside") %>%
    group_by(StimSet, Target) %>%
    summarise(n = n()) %>%
    mutate(freq = n / sum(n))

  # need to add "space" as level of Target in human data
  ch_cond$Target <- factor(ch_cond$Target, levels = c(shapes_vec))
  fig_targ_hit <- ggplot(ch_cond, aes(x=StimSet, y=freq, shape=Target, color=Target)) + # , size=n
    geom_point(alpha=.8) +
    scale_shape_manual(values=10:20) +
    #geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    xlab("Stimulus Set") + ylab("Rate of Target Object Hit") + ylim(0,ymax) + 
    langcog::scale_fill_solarized() + ggthemes::theme_few() + 
    geom_hline(yintercept=.33, lty='dashed')
}
#print(fig_targ_hit)
fig_targ_hit <- target_hit_plot(hum, ymax=.66, shapes)
```

```{r target-space, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.4, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Relative frequency of targeted object as a function of stimulus set."}

ch_cond <- hum %>% 
  filter(is.element(TargetSpace, shapes)) %>%
  group_by(StimSet, TargetSpace) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

ch_cond$TargetSpace <- factor(ch_cond$TargetSpace, levels = shapes)
fig_targ_space <- ggplot(ch_cond, aes(x=StimSet, y=freq, shape=TargetSpace, color=TargetSpace)) + # , size=n
  geom_point(alpha=.8) +
  scale_shape_manual(values=10:19) +
  #geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  xlab("Stimulus Set") + ylab("Rate of Targeted Location") + ylim(0,.66) + 
  langcog::scale_fill_solarized() + ggthemes::theme_few() + 
  geom_hline(yintercept=.33, lty='dashed')
#print(fig_targ_space)
```

```{r combined-fig, fig.env = "figure*", fig.pos = "h", fig.align="center", fig.width=7, fig.height=3.2, set.cap.width=T, num.cols.cap=2, fig.cap = "Relative frequency of objects 1) chosen as objects for dropping, 2) hit as targets (pink dots represent hitting empty space), and 3) whose drop location was closest to a target object, as a function of target/drop set pairing."}

ggarrange(fig_drop, fig_targ_hit, fig_targ_space, nrow=1, 
          #labels = c("Drop Choice", "Targeted Object Hit", "Targeted Location"),
          common.legend=T, legend="bottom")
```

@@

```{r unique-drop-objects, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Proportion of unique objects selected by participants. Error bars show boostrapped 95\\% confidence intervals."}
# 12 trials, 9 objects; random selection would be 9/12 = 75%...
uniq_s <- hum %>% group_by(SID, Age_Group) %>%
  summarise(DropObjs = n_distinct(DropChoice), N=n(), prop=DropObjs/N)

@@

```{r drop-by-target, include=F, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Heatmap of participants' drop object choices by target location."}
ch_loc <- hum %>% filter(DropLocation!="outside" & DropLocation!="space") %>%
  group_by(DropChoice, DropLocation) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(ch_loc, aes(x=DropLocation, y=DropChoice)) + geom_tile(aes(fill=freq)) + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  xlab("Target Location") + ylab("Drop Object") +
  ggthemes::theme_few()
```

### Tower Task Results

```{r load-tower-data}

tower = read.csv2("../analysis/human-data/curiobaby_drop-data - cool tower.csv", sep=',', header=T, stringsAsFactors=F)
hum = subset(tower, exclude!="Y")

#length(unique(tower$SID)) #61
Nsubj_tower = length(unique(hum$SID)) # 51 subjects

hum$age_group = as.numeric(hum$age_group)
hum$age_rounded = as.numeric(hum$age_rounded)
hum$height_tallest = as.numeric(hum$height_tallest)
hum$tallest_objects = as.character(hum$tallest_objects)
hum$base_tallest = as.character(hum$base_tallest)

ht_age = cor.test(hum$age_rounded, hum$height_tallest)
```

The average height of the constructed towers was `r round(mean(na.omit(hum$height_tallest)), 1)` blocks.
We examined tower height as a function of age, and found that older children tend to build higher towers (`r cor_apa(ht_age, format="rmarkdown", print=F)`).

```{r plot-height-by-age, include=F, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Tower height by age."}
ch_height <- hum %>% group_by(age_rounded, height_tallest) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(ch_height, aes(x=age_rounded, y=height_tallest)) +
  geom_smooth(method="lm", level=0.90) + geom_point(alpha=.8) +
  scale_shape_manual(values=10:19) +
  xlab("Age") + ylab("Height_Tallest") + ylim(0,10) +
  langcog::scale_fill_solarized() + ggthemes::theme_few()
```

Next we examine the number of times participants supported each object with each other object. 
```{r heatmap-support-by-supported-counts, include=T, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Counts of support and supported object relations in participants' towers."}
tower = subset(tower, tallest_objects!="" & tallest_objects!="pentagon+octahedron, trig prism") # 5 kids who arranged horizontally and 1 kid chose 2 - exclude
df = data.frame(support=NA, supported=NA, n=NA)
for(s in 1:nrow(tower)) {
  objs = unlist(str_split(tower[s,]$tallest_objects, ", "))
  #print(objs)
  for(i in 2:length(objs)) {
    df = rbind(df, cbind(support=objs[i-1], supported=objs[i], n=1))
  }
}
df = na.omit(df)
hm <- df %>% group_by(support, supported) %>% count(N=n()) 
hm %>% ggplot(aes(x=support, y=supported)) + geom_tile(aes(fill=N)) + geom_text(aes(label = N)) +
  scale_fill_gradient(low = "white", high = "steelblue") + ggthemes::theme_few()
```

(Could instead show P(supported object | support object).)
```{r heatmap-support-by-supported-probs, include=F, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Conditional probabilities of supported object given support object in participants' towers."}
tower = subset(tower, tallest_objects!="" & tallest_objects!="pentagon+octahedron, trig prism") # 5 kids who arranged horizontally and 1 kid chose 2 - exclude
df = data.frame(support=NA, supported=NA, n=NA)
for(s in 1:nrow(tower)) {
  objs = unlist(str_split(tower[s,]$tallest_objects, ", "))
  print(objs)
  for(i in 2:length(objs)) {
    df = rbind(df, cbind(support=objs[i-1], supported=objs[i], n=1))
  }
}
df = na.omit(df)
hm1 <- df %>% 
  group_by(support, supported) %>% 
  summarise(n = n()) %>% 
  group_by(support) %>%
  mutate(freq = n / sum(n))
hm1 %>% ggplot(aes(x=support, y=supported)) + geom_tile(aes(fill=freq)) + geom_text(aes(label = round(freq, 2))) +
  scale_fill_gradient(low = "white", high = "steelblue") + ggthemes::theme_few()
```

<!-- Finally, we look at the distribution of tower heights as a function of the objects chosen as the base of the tower. -->

```{r base-and-top-objects, include=F, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Distribution of base and top objects in participants' towers."}
ch_extremes <- hum %>% 
  group_by(base_tallest) %>% 
  count(N=n()) %>%
  filter(base_tallest!="" & base_tallest!="pentagon+octahedron")

>>>>

---
title: "What if you drop that? Developmental changes in curiosity about physical interactions"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George (kachergis@stanford.edu)} \\ Department of Psychology, Stanford University \\ Madison, WI 53706 USA
    \AND {\large \bf Judith Fan (SDJ@Macc.Wisc.Edu)} \\ Department of Psychology, University of San Diego \\ Madison, WI 53706 USA}

abstract: >
     Curiosity is a fundamental driver of human behavior, and yet because of its open-ended nature 
     and the wide variety of behaviors it inspires in different contexts, it is remarkably difficult 
     to study in a laboratory context. A promising approach to developing and testing theories of 
     curiosity is to instantiate them in artificial agents that are able to act and explore in a 
     simulated environment, and then compare the behavior of these agents to humans exploring the 
     same stimuli. Here we propose a new experimental paradigm for examining children's--and curious
     AI agents'--curiosity about objects' physical interactions, using a task that is both open-ended
     enough to allow room for curiosity, but also constrained enough to make detailed behavioral 
     comparisons. We children's choices (N=XX; 2-6 years of age) to those of adults (N=XX), and find
     systematic preference for drops affording particular relations (e.g., containment)

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
```

```{r, load-data}
# preprocessed in analysis/exp2_analysis.Rmd
load(file=here("paper/processed_data.RData"))

Nch = length(unique(child_long$ID)) # 62
Nad = length(unique(adult_long$ResponseId)) # 35

age <- child_long %>% distinct(ID, Age) %>% group_by(Age) %>%
  summarise(Age = mean(Age), N= n())
```

@@

<!-- Curiosity is important... end with core question about what gives rise to curiosity. -->
Curiosity is a hallmark aspect of human intelligence. 
From infants exploring the objects in their environment to scientists exploring the frontiers of our solar system, humans are highly motivated to seek out new knowledge and experiences. 
However, although such exploratory behavior has long been recognized as a critical component of human learning [@james1983talks] and cognitive development [@Gopnik2009; @Piaget1952], formal theories that explain human curiosity and how it drives exploratory behavior have remained elusive [@Kidd2015]. 
Moreover, extant theories have rarely provided quantitatively precise enough predictions to be directly compared to empirical measurements of curiosity-driven behavior in humans.

@@

Our computational theory is inspired by classic descriptive theories that posited that curiosity-driven behavior was the consequence of attentional capture by novel stimuli [e.g., @Berlyne1954; @Fantz1964], as well as theories providing qualitative explanations for curiosity-driven behavior as the result of learners' preferences for stimuli that are "moderately discrepant" in relation to their current knowledge state, and thereby provide opportunities to learn [@Kinney1976]. 
Recent studies have provided empirical support for this basic idea: XX-old infants are sensitive to stimulus complexity and prefer to look at moderately complex stimuli rather than overly simple or overly complex stimuli; @Kidd2012; @Kidd2014].
According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957].
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.  
Our approach leverages theory and insights from reinforcement learning (RL), including the value of mechanisms that equip RL agents with "intrinsic motivation" — a drive to explore the state space even when rewards are sparse or absent, and favor actions with uncertain outcomes, because they may result in the discovery of new policies with high expected values [@Schmidhuber2010].
Our modeling approach is resonant with recent work that has instantiated such intrinsic-motivation mechanisms in robots to help them learn robust ways to predict physical events in the world world [@Oudeyer2007; @Oudeyer2013], although this work typically involves pretraining models on a separate physical prediction task before implementing curiosity-driven learning, and has not yet been directly compared to human behavior in the same prediction tasks.

Our empirical study is motivated by a large body of prior work in developmental psychology investigating the development of knowledge about physical objects, their properties, and how they interact (CITE), thus making exploration of physical objects a natural choice of domain to explore the implications of our theory. 
However, the measures used in this literature have typically been coarse-grained with respect to infant's and children's physical knowledge, and thus do not specify or constrain which events might moderately discrepant from this knowledge. 
For instance, a typical measure of physical curiosity has been longer looking times to surprising events than to expected events (e.g., objects appearing to pass through a wall; @Stahl2015; @Baillargeon2007). 
Other work that has investigated exploratory behavior in children have used more granular measures, such as counting the number of functions discovered while playing with novel and complex toy [e.g., @Cook2011; @Bonawitz2012; @Gweon2014]. 
Moreover, these studies have yet to directly constrain theories that predict which functions children will discover, and in what sequence. 
What both of the standard looking-time and toy-exploration measures fail to capture when characterizing children's curiosity are the ways in which children actively learn by intervening on the world and observing the consequences of their actions [@Gopnik2009; @gureckis2012self]. 
To address these limitations, we developed a novel physical exploration task in which children, adults, or AI agents choose which series of physical experiments to perform and observe the results of each in real time.

@@

# Experiment 1: Adults

@@

### Participants
Participants were `r Nad` adults recruited online via Amazon Mechanical Turk.

@@

```{r adult-agg}
adult_trial_agg <- adult_long %>% group_by(drop, target) %>% # alt, 
  summarize(chose_target=mean(chose_target, na.rm=T), n=n())
```

# Experiment 2: Children

### Participants
Participants were `r Nch` children recruited online via XX.
Participant exclusions were made based on cases where i) the participant did not complete more than half of the study play session or ii) the parent did not consent for video recording of study. 
After exclusions, results from `r Nch` were analyzed, including `r age[1,]$N` 3-year-olds, `r age[2,]$N` 4-year-olds, `r age[3,]$N` 5-year-olds, `r age[4,]$N` 6-year-olds, and `r age[5,]$N` 7-year-olds.

@@

```{r fig1-sets, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.25, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "Set of 3D objects used for dropping and as targets. [REPLACE]"}
img <- png::readPNG("figs/drop_sets.png")
grid::grid.raster(img)
```

@@

After the parent provided informed consent, children were assigned to Order 1 or Order 2. Participants were assigned to condition in counterbalanced order.
We introduced children to a set of 3D-printed toy objects ("blocks").

An example trial is illustrated in Figure 2.

Based on piloting, we estimated the activity would would only require X minutes to complete. 
In both conditions, a video camera was used to record the play session from an angle above the bin, to show child's block selection and drop location as well as child's completed tower. 
After child notified the researcher that they were finished building their tower, the session was completed.

```{r fig2-task, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Example trial in which the participant chose to drop the pyramid on the pentagonal prism. [REPLACE]"}
img <- png::readPNG("figs/DropTaskFigure.png")
grid::grid.raster(img)
```

@@

```{r child-data-agg}
child_trial_agg <- child_long %>% group_by(relation, drop, target) %>% # alt, 
  summarize(chose_target=mean(chose_target, na.rm=T), n=n())

ch_ad_cor <- cor.test(child_trial_agg$chose_target, adult_trial_agg$chose_target)

trial_tab <- child_trial_agg %>% select(-n) %>% 
  bind_cols(adult_target_choice = adult_trial_agg$chose_target) %>% 
  rename(child_target_choice = chose_target) 
```

We first examine children's choice of drop objects and target objects to determine if children were choosing randomly, or had consistent preferences for some objects.

Were children's drop choices consistent with adults'?
Correlation between children's individual proportion of "missed" targets and their age (`r cor_apa(ch_ad_cor, format="rmarkdown", print=F)`).
Table 2 shows a summary of participants' choice of target objects per set.

<!-- Finally, we examine a more lenient analysis children's choice of target location (i.e., the identity of the closest target), regardless of whether there was a collision of drop and target objects. -->
<!-- Chi-square tests of children's target collisions once more found that ... -->
<!-- Table 3 shows a summary of participants' choice of target locations per set. -->

<!-- In summary, children showed some consistent preferences in their choice of objects to drop from each set, but no systematic preference for targeting particular objects. -->
<!-- We now investigate whether there were interactions of drop objects and targets that children found particularly appealing. [ToDo: big chisq test?] -->

<!-- Finally, we analyze how exploratory children were in their responses: across the 12 trials, what proportion of the 9 available objects did each child utilize as drop objects? What proportion of the objects were targeted? -->

@@

#fig_drop <- drop_choice_plot(hum, shapes)
```

@@

```{r unique-drop-objects, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Proportion of unique objects selected by participants. Error bars show boostrapped 95\\% confidence intervals.", eval=F}
# 12 trials, 9 objects; random selection would be 9/12 = 75%...
uniq_s <- hum %>% group_by(SID, Age_Group) %>%
  summarise(DropObjs = n_distinct(DropChoice), N=n(), prop=DropObjs/N)

@@
